Inter-type declarations (ITDs) provide a powerful yet simple modularization
mechanism. The possibility to extend existing classes modularly without
ahead-of-time planning is not only useful to separate different concerns
but also extremely convenient for modular extensibility when software
evolves.

A common criticism of ITDs is their global scope which arguably leads to
poor information hiding, a topic that has gained renewed interest with the 
emerging support for modules in Java 7. Another disadvantage is that the 
class hierarchy is destructively updated, preventing multiple variants of 
classes with different sets of ITDs applied. It is worth noting that 
this drawback is not shared by more traditional extension mechanisms such 
as visitors. The problem of how to handle multiple variants of the same 
class exist in plain Java as JAR hell, and there has been a wealth of work 
on module systems for Java-like languages to improve on status quo. We use 
this existing work on Java modules in an attempt to solve the class 
variant problem for ITDs.

The emerging support for modules in Java 7 enhances information hiding and
extended module proposals such as Strni\v{s}a gives hope for simultaneous
deployment of multiple versions of the same library in different modules.
Modules provide information hiding at a level higher than packages. Module
systems like OSGi bundles\cite{OSGi4} and the proposed Java module system\cite{JSR277}
allow the explicit definition of the dependencies of a module. We can extend
the information hiding features of these module systems to extend to aspects, 
so as to limit the scope of ITDs.
Another module system, iJAM \cite{iJAM}, adds explicit module instantiation, 
which allows multiple versions of the same module to coexist in a single compilation.
This is built upon by some of our previous work \cite{modulesastypes}, which 
adds the idea of treating modules as object-oriented types, with instantiation
similar to iJAM and additional operations to allow explicit and fine-grained
module instance sharing. 

%ITDs are global in their nature which makes local
%reasoning somewhat difficult, and as an extensibility mechanism they can be
%improved by enabling deployment of multiple versions of libraries, 
%each woven a with different set of ITDs.


Previous work show how aspects can be improved using modules for point-cut
and advice \cite{openmodulesaj}. Aspects don't work very well without 
modules, due to global scope and implicit dependencies. We believe that 
such benefits are even more important for a system with inter-type 
declarations. In this paper we therefore present a module system that 
supports inter-type declarations and improve their use when extending a 
system in a modular fashion.

%Aspect instantiation has always been a sticky subject (e.g. doubly applied
%pointcuts in AspectJ abstract aspects)

We have extended previous work on object-oriented modules for Java \cite{modulesastypes}, 
and implemented the proposed module system as an extension to the 
Jast\-Add system, an aspect oriented compiler construction system 
extending Java with support for attribute grammars and ITDs. That system is itself an extension 
to the Jast\-Add Extensible Java compiler (Jast\-AddJ), a Java frontend and compiler
which is used in for instance the AspectBench compiler for AspectJ, and the 
Soot bytecode manipulation framework. The implementation of ITDs in the JastAdd
system is actually shared with the implementation of ITDs in the
AspectBench Compiler for AspectJ. 

To evaluate the module system we performed a case study where JastAddJ was 
refactored to use the proposed module system. It is important to 
distinguish JastAdd, a meta compiler system used to build JastAddJ, from JastAddJ, a Java 
compiler, since both are used as examples in the paper.

In our previous work we explain how to use ITDs as one of the main
modularization mechanisms when building extensible compilers \cite{aosd08abc}. Our current
work involves using the same techniques to generate IDEs for a wide range
of dialects of Java. In such an IDE we may for instance want to use numerous variants of the
same frontend that slightly differ to support different dialects. We also
want to use a pure frontend for error checking while a backend is also
needed to support code generation.
That work highlights some deficiencies to ITDs from an extensibility point of
view compared to the more traditional use of visitors.
Inter-type declarations have the advantages of not requiring ahead-of-time 
planning, the ability to add state, no need for boiler-plate code, and 
being less error prone since adherence to framework 
conventions to enable dispatch is not necessary.
However, ITDs do not provide the same level of extensibility as visitors.
ITDs are currently destructive updates of the class hierarchy. The base version and the
extended version can not co-exist. This is the main motivation for our
work. 
Our case study shows that the proposed module system solves these deficiencies completely in an
elegant fashion, and also enhances information hiding by restricting 
visibility of large parts of the system that used to be globally visible.

These are the main contributions of this paper:
\begin{itemize}
\item The design of a module system for ITDs that improves information
hiding and extensibility.
\item An implementation as a modular extension to the Jast\-Add meta 
compiler system.
\item A case study where implementation of the Java compiler JastAddJ is retrofitted to use the
proposed module system.
\end{itemize}

The rest of this paper is structured as follows: In
Section~\ref{section:itdvisitors} we present a detailed example that
highlight the merits and deficiencies of ITDs compared to a visitor based
approach. A module system that shows how that example can be improved is
presented in Section~\ref{section:jastaddmodules} and we present a case
study where an extensible Java compiler is retrofitted to use that module
system in Section~\ref{section:casestudy} where we also discuss 
the advantages it brings compared to the original implementation. A
brief overview of the module system implementation is presented in
Section~\ref{section:implementation} and we discuss related work in
Section~\ref{section:related} and conclude in
Section~\ref{section:conclusions}.

