Implement as a class loader. This will probably bring in a lot of issues related to
lookup once the class is loaded. This (might) work if the JVM is modified to
be aware of module qualifiers (fat chance, but who knows?).

Currently, module references are dynamically typed. Their types change if they
are merged or are replaced with other modules. See if static typing of the 
references would make more sense, or would just be too restrictive.

Explore looser merge that synthesizes a new module if there are no conflicts
(but there are problems here due to possibility of introducing class lookup ambiguity
on existing modules).

Explore recursive merge that actually does merging instead of just creating a
new instance. Will this break other modules?

Weaken merge/replace inheritance. Allow a super() call and the option not to use it, 
but guarantee that the signature of the module is not changed so as not to break
the module's clients

More operations on modules than merge and replace. What about global operations? (e.g.
change all references to this module/module interface to point to this new module)

Packages are still just class modifiers. Find a way to integrate packages in the type
system itself. A possibly useful addition to \texttt{package-info.java} is the specification
of required classes as annotations

\begin{lstlisting}
//file package-info.java
package org.x.y.lib;

@RequiredClass("Main");
@RequiredClass("Parser");
@RequiredClass("Scanner");
\end{lstlisting}

The annotations would cause an error in a build if one of the required classes were not
found in the package. This would reduce the problems caused by split packages, and it should
also be possible to use these in a manner similar to module interfaces to specify the kind
of package that an application requires.

The type system is almost a copy of every object-oriented type system. Find more
operations/relations that are inspired by this similarity.

Implement the module qualified names in reflection.

Find ways to make the merge/replace (or even import) operations possible at run time.
It would be great to replace an existing version of a module with a dynamically generated
extended version at run time.